Quantifying and Modeling Global Water Accessibility including Water Typology

### 2021-01-25 docs/manuscript/manuscript.tex
Revision.

\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting
impact on MDG progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
water quality and sanitary risk utilizing national water quality data. 
%\citet{onda2012global} used principal components
%analysis to analyze the national environment and development indicators and then created models. 
Overall, their analysis highlights the potential fatigue illness caused by people transporting the water if it is done manually. How the
modes of transportation can affect one's health.

\citet{sima2013water} highlights the relation of the topic where it discovered that over 76 percent of the city's water
is utilized by less than ten percent of household piped water.  The study is focused in Kisumu Kenya and Sima
interviewed 260 informal water business operators. This article revealed a new insight where the majority of the
population in Kisumu relied on purchased water from kiosks (1.5 million $m^{3}$ per day) and used hand-drawn water-carts
(0.75 million $m^{3}$ per day)\citep{sima2013water}. Also, we discovered that the water trucking industry utilized the
most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35
$MJ/m^{3}$)\citep{sima2013water}. The article utilized SPSS Version 19.0 to analyze the data.

\citet{ho2014challenge} highlights a head-to-head comparison of such indicators with other possible distance and time
metrics among rural 1,103 households in Nampula province, Mozambique. The researchers found out that Euclidean line
distance to be a good representative for route distance ($R^{2} = 0.98$), while self-reported travel time is a poor
representative ($R^{2} = 0.12$). One of key insights was that, 15-minute decrease in one-way walk time to water source
is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child
mortality rate \citep{ho2014challenge}. This research can be a helpful tool to understand how long people are traveling
in a rural area in Africa where many of our data focus. Also, understanding how people are traveling to receive the
water can better understand the condition of the rural people and implement better policies that help them to thrive.

\citet{bartram2014globala} highlights a background information about the Millennium Development Goals. How it has
established global targets for drinking water and sanitation access (Bartram et al. 2014). The article indirectly gives
much information about progress towards focused targets, directed by international monitoring, has reduced the global
disease burden and improved the quality of life \citep{bartram2014globala}. It utilizes a variety of sources such as
JMP, DHS, MICS surveys to analyze the data of many different locations in the world, therefore, it is a broad scope of
study of the progress of accessible and drinking water and sanitation globally. Our research can study further what
modes of transportation of clean drinking water would affect the progress of improving drinking water and sanitation in
one's community. Also, understanding how long it takes to attain the water can improve the quality of the services
provided for the people.

\citet{onda2014country} proposed a country typology based on water and sanitation indicators. They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people. This work improves on previous geography or income-based country clustering. 
%Their research can improvise upon their finding and focus on identifying which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.

>>>>

\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting
impact on (MDGs) progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
water quality and sanitary risk utilizing national water quality data. 
Overall, their analysis highlights the potential fatigue can have causes that aren't due to underlying disease. The main example would be heavy exertion of carrying water.  Improvement of modes of transportation can enhance water accessibility which can improve one's health overall. 
modes of transportation can affect one's health.

\citet{sima2013water} highlights the relation of the topic where it discovered that over 76 percent of the city's water
is utilized by less than ten percent of household piped water.  The study is focused on Kisumu Kenya and Sima
interviewed 260 informal water business operators. This article revealed a new insight where the majority of the
population in Kisumu relied on purchased water from kiosks (1.5 million $m^{3}$ per day) and used hand-drawn water-carts
(0.75 million $m^{3}$ per day)\citep{sima2013water}. Also, we discovered that the water trucking industry utilized the
most energy consumption and was high cost: water delivery trucks have the highest per cubic meter energy demand (35
$MJ/m^{3}$)\citep{sima2013water}.

\citet{ho2014challenge} investigates a head-to-head comparison of such indicators with other possible distance and time
metrics among rural 1,103 households in Nampula province, Mozambique. The researchers found out that Euclidean line
distance to be a good representative for route distance ($R^{2} = 0.98$), while self-reported travel time is a poor
representative ($R^{2} = 0.12$). One of the key insights was that a 15-minute decrease in one-way walk time to a water source
is related to a 41\% reduction in diarrhea, improved children's nutrition, and an 11\% reduction in under-five child
mortality rate \citep{ho2014challenge}.

\citet{bartram2014globala} highlights background information about the Millennium Development Goals. How it has
established global targets for drinking water and sanitation access (Bartram et al. 2014). The article indirectly gives
much information about progress towards focused targets, directed by international monitoring, has reduced the global
disease burden and improved the quality of life \citep{bartram2014globala}. It utilizes a variety of sources such as
JMP, DHS, MICS surveys to analyze the data of many different locations in the world, therefore, it is a broad scope of
study of the progress of accessible and drinking water and sanitation globally. Our research can study further what
modes of transportation of clean drinking water would affect the progress of improving drinking water and sanitation in
one's community.

\citet{onda2014country} proposed a country typology based on nine water and sanitation indicators. They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people (93\%). This work improves on previous geography or income-based country clustering.
